---
title: "Reproducible Research: Peer Assessment 1"
author: "Gerard Lopez"
date: "Sunday, November 16, 2014"
output: 
  html_document:
    keep_md: true
---


```{r setoptions, echo=TRUE}
opts_chunk$set(echo=FALSE)
```

## Loading and preprocessing the data

```{r}
library("lattice")
library("lubridate")
library("ggplot2" )

rm(StepsByDay, StepsByDayMean, StepsByDayMedian, StepsByInterval, timeInterval)

unzip("activity.zip", overwrite=TRUE)
stepData <- read.csv("activity.csv", colClasses=c('integer','character', 'integer'))
stepData$timeInterval <- as.POSIXct(strptime(formatC(stepData$interval, width = 4, format="d", flag ="0"), "%H%M"))

```

## What is mean total number of steps taken per day?
1. Make a histogram of the total number of steps taken each day

```{r}
StepsByDay <- by(stepData$steps, stepData$date, FUN=sum)

hist(StepsByDay)
```

2. Calculate and report the mean and median total number of steps taken per day

```{r}
StepsByDayMean <- mean(StepsByDay, na.rm = TRUE)
StepsByDayMedian <- median(StepsByDay, na.rm = TRUE)

```


## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}

StepsByInterval<- aggregate(stepData$steps, list(timeInterval=stepData$timeInterval), mean, na.rm= TRUE)

#set x limits
beginning <- format(as.POSIXct(paste(Sys.Date(),"00:00")), format="%Y-%m-%d %H:%M")
ending <- format((as.POSIXct(paste(Sys.Date(),"00:00")) + 24*60*60), format="%Y-%m-%d %H:%M")

xyplot( x ~ timeInterval, data=StepsByInterval, type = "l"
        , xlab = "Time", ylab = "Average # of Steps"
        , xlim=c(as.POSIXct(beginning),as.POSIXct(ending))
        , ylim=c(0,225)
        , scales = list(tck=c(1,0),
                        x=list(tick.number = 24, cex=0.8, rot=90, format="%H:%M"),
                        y=list(tick.number = 5))
        , main = "Average Number of Steps During 5 Minute Intervals Over 24 Hours"
        )

#
```



2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?


```{r}
temp <- StepsByInterval[ StepsByInterval$x == max(StepsByInterval$x, na.rm=TRUE), ]
MaxStepInterval = strftime(temp$timeInterval, "%H:%M")

rm(temp)


```




## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
rowsWithNAs <- nrow(stepData[!complete.cases(stepData),])
```


2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r}

One option:

# For each thing, make some kind of aggregate calculation
        #StepsByInterval already does this, but has interval as a time...need to put it back as it was.
temp <-  StepsByInterval
temp$new <- as.integer(strftime(temp$interval, "%H%M") )




        
# Merge the aggregate number back into the original data as a new imputation column based on a link between individual thing instances and the aggregate thing instance

ImputedData <- merge(stepData, StepsByInterval, by = c("timeInterval"))
ImputedData$steps = ifelse(is.na(ImputedData$steps), ImputedData$x, ImputedData$steps)
# ImputedData <- ImputedData[order(ImputedData$date,ImputedData$interval),]


# Get rid of the imputation column to tidy up.
ImputedData$x <- NULL 


```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r}

```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


```{r}

```




## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.


```{r}

```


2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}

```

